---
title: "Activity Tidy Data - Cookbook"
author: "K Samanta"
date: "26 October 2015"
output: md_document
---

# Activity Tidy data cookbook

This document provides the definition of the activity data summaraised at the Subject and Activity level. The data structure and definition are explained at the end of processing.  

## Download Files  

```{r}
# library(utils)
# #Download File
# setwd("C:/CourseraR/GettingandCleaning/Assigment1/code")
# fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
# download.file(fileURL, destfile = "../data/FUCR.zip")
# unzip("../data/FUCR.zip", exdir = "../data")

```
## Load Training Data  
```{r}

#Load Training DataSets
actv_lvls <- read.table(file = "../data/UCI HAR Dataset/activity_labels.txt")
feat <- read.table(file = "../data/UCI HAR Dataset/features.txt")

sub_trn <- read.table("../data/UCI HAR Dataset/train/subject_train.txt")
X_trn <- read.table("../data/UCI HAR Dataset/train/X_train.txt")
Y_trn <- read.table("../data/UCI HAR Dataset/train/y_train.txt")

names(X_trn) <- feat$V2
Y_trn$V1 <- factor(Y_trn$V1, labels = actv_lvls$V2)
names(Y_trn) <- "Activity"
sub_trn$V1 <- factor(sub_trn$V1)
names(sub_trn) <- "Subject"

all_trn <- cbind(sub_trn, Y_trn, X_trn)
```
## Load Test Data  
```{r}
#Load Test DataSets
sub_test <- read.table("../data/UCI HAR Dataset/test/subject_test.txt")
X_test <- read.table("../data/UCI HAR Dataset/test/X_test.txt")
Y_test <- read.table("../data/UCI HAR Dataset/test/y_test.txt")

names(X_test) <- feat$V2
Y_test$V1 <- factor(Y_test$V1, labels = actv_lvls$V2)
names(Y_test) <- "Activity"
sub_test$V1 <- factor(sub_test$V1)
names(sub_test) <- "Subject"
all_test <- cbind(sub_test, Y_test, X_test)
```
## Merge Training and Test data
```{r}
#Combine training and test data set
all_data <- rbind(all_test, all_trn)
```
## Subset for mean and standard deviation features 
```{r}
#Identify subset of variables for mean and standard deviation
feat_mn <- feat[grep("mean", feat$V2), 1 ]
feat_std <- feat[grep("std", feat$V2), 1 ]
feat_mn_std <- c(1,2,(feat_mn+2), (feat_std+2))
all_data_mn_std <- all_data[,feat_mn_std]
```
## Summarise by Subject and Activity
```{r}
#Summarise data and create the tidy text file
tidy_data <- aggregate(all_data_mn_std[-c(1:2)], 
                       by = list(all_data_mn_std$Subject, all_data_mn_std$Activity),
                       FUN = mean)
names(tidy_data)[1] <- "Subject"
names(tidy_data)[2] <- "Activity"
names(tidy_data) <- gsub("-","",names(tidy_data))
names(tidy_data) <- gsub("\\(\\)","",names(tidy_data))
names(tidy_data) <- gsub(",","",names(tidy_data))

write.table(tidy_data, file = "tidy_data.txt", row.name=FALSE )

```
The file contains the following features for mean and standard deviation -  
```{r}
names(tidy_data)
```
